From noise to knowledge: how randomness generates novel phenomena and information

### 2017-10-21 code/appendix.Rmd
WIP

p <- list(r = .4, K = 1, sigma = .2, N = 500)

@@

@@

```{r}
(p$K + sqrt(p$K^2 - 8 * p$sigma^2) )/ 2
```

@@

@@

Understanding the origin of sustained oscillations has long been a central question in ecology [@Hastings]

@@

df <- p %>% rowwise() %>% do(f(.))

```{r}
df %>%
  gather(species, pop, -t, -sigma) %>%
  ggplot(aes(t, pop, col=species)) + geom_line() + facet_wrap(~sigma)
```

@@

@@

```{r}
system.time({
may  <- nimble::nimbleCode({

@@

model <- nimbleModel(may,constants = p, inits = list(x0 = 1.2))
cmodel <- compileNimble(model)
set.seed(123)

simulate(cmodel)
df <- tibble(t = seq_along(cmodel$x), x = cmodel$x)
})

df  %>% ggplot(aes(t,x)) + geom_line()
```

@@

model <- nimbleModel(tip,constants = p, inits = list(x0 = 1))
cmodel <- compileNimble(model)
set.seed(123)

simulate(cmodel)
df <- tibble(t = seq_along(cmodel$x), x = cmodel$x)

```{r}
df %>% ggplot(aes(t,x)) + geom_line()
```

>>>>

p <- list(r = .4, K = 1, sigma = .15, N = 500)

@@

df %>% 
  filter(x > 0) %>% 
  group_by(t) %>%
  summarise(ave = mean(x), sd = sd(x)) %>%
  write_csv("inflation.csv")

@@

Analytical predicted population size, as a fraction of $K$, is:

```{r}
(1 + sqrt(1 - 8 * p$sigma^2/p$K^2) )/ 2
```
Note this is independent of $r$, and increases with $\sigma/K$, the larger the noise $\sigma_g$ as a fraction of the carrying capacity $K$.  Population sizes can also be inflated relative to their deterministic equilibrium whenever the second derivative is positive (as in the lower equilibrium in the May model of alternative stable states, considered below.)  Stronger nonlinearities (in particular, a larger second derivative at the equilibrium) can drive larger inflationary effects.

@@

```{r}
p <- list(r = .5, K = 2.1, Q = 5, H = .4, sigma = .04, a = 0.27, N = 1e4)

x <- seq(0,2, length.out = 100)
df <- data.frame(x, f = f(x), g = g(x))

ggplot(df, aes(x)) + geom_line(aes(y = f)) + geom_line(aes(y = g), col="red")

ggplot(df, aes(x)) + geom_line(aes(y = f-g)) + 
  geom_hline(aes(yintercept=0), lty=2) +
  coord_cartesian(xlim = c(0,1.5), ylim=c(-.1,.1))  
```

@@

Understanding the origin of sustained oscillations has long been a central question in ecology [@Hastings1996]

@@

quasicycle_df <- p %>% rowwise() %>% do(f(.))        
write_csv(quasicycle_df, "quasicycles.csv")

quasicycle_df %>%
  gather(species, pop, -t, -sigma) %>%
  ggplot(aes(t, pop, col=species)) + geom_line() + facet_wrap(~sigma)
```

@@

May's model

@@

may  <- nimble::nimbleCode({

@@

model <- nimbleModel(may,constants = p, inits = list(x0 = 1.2))
cmodel <- compileNimble(model)
set.seed(123)
simulate(cmodel)

tibble(t = seq_along(cmodel$x), x = cmodel$x) %>% write_csv("noisy_switch.csv")
read_csv("noisy_switch.csv")  %>% ggplot(aes(t,x)) + geom_line()
```

@@

model <- nimbleModel(tip,constants = p, inits = list(x0 = 1))
cmodel <- compileNimble(model)
set.seed(123)
simulate(cmodel)

tibble(t = seq_along(cmodel$x), x = cmodel$x) %>% write_csv("tipping.csv")

read_csv("tipping.csv") %>% ggplot(aes(t,x)) + geom_line()
```

### 2017-10-21 paper.Rmd
WIP

@@

>>>>

**Inflation**

@@

Understanding the origin of sustained oscillations has long been a central question in ecology [@Hastings1996].  The simple Lotka Volterra model produces only neutrally stable centers, while the more realistic model that introduces carrying capacity for prey as they compete for resources results in a model that produces damped oscillations, eventually returning to a stable
